---
title: "envs193ds_homework_3"
format: html
---

## Part 1. Set up tasks

\[GitHub Repository\] (https://github.com/norsmith025/ENVS-193DS_homework-03.git)

```{r}
#| message: false # hide messages in render

# read in packages
library(tidyverse) # load tidyverse 
library(here) # load here
library(gt) # load gt
library(janitor) # load janitor
library(readxl) # load readxl

```

## Part 2. Problems

#### Problem 1. Personal data

a.  I can sort my hydration levels into 2 groups, Low Water Intake and High Water Intake, and find the mean productivity level, from a scale of 1-5, for each group. This would be informative because I think I am *generally* more productive on days where I am more hydrated.

b.  Visualization

    ```{r}
    # create "my_data" object from "my_data.csv" from "data" folder
    my_data <- read_csv("data/my_data.csv", skip = 1) 

    # clean data and create columns
    my_data <- my_data %>%
      mutate(Water = as.numeric(`Water (oz)`), # convert to numeric values and store in column "Water"
             Productivity = as.numeric(`Productivity (1-5)`), # convert to numeric values and store in column "Productivity"
             Hydration_Level = if_else(Water >= 32, "High Water Intake", "Low Water Intake")) # create new column "Hydration_Level" based on Water with threshold of 32 oz

    # create the plot
    ggplot(my_data, aes(x = Hydration_Level, y = Productivity, color = Hydration_Level)) +
      geom_jitter(width = 0.3, alpha = 0.8) + # add jittered points 
      stat_summary(fun = mean, geom = "crossbar", width = 0.5, color = "darkgrey") + # add horizontal bar to show mean productivity for both hydration groups
      scale_color_manual(values = c("High Water Intake" = "cadetblue", "Low Water Intake" = "salmon")) + # manually set colors for both hydration groups
      labs(x = "Hydration Level", # label x-axis
           y = "Productivity (1-5)", # label y-axis
           title = "Impact of Hydration on Productivity Levels", # create title
           caption = "Productivity levels grouped by Low and High Water Intake (Hydration Levels).\nMy mean productivity is higher on days with ≥32 ounces (oz) of water intake,\n which indicates a possible positive") + # add caption
      theme_minimal() + # use minimal theme 
      theme(plot.caption = element_text(hjust = 0.5,
                                        size = 10, 
                                        margin = margin(t = 10))) + # align caption 
      theme(legend.position = "none") # hide legend 

    ```

c.  Add caption: Productivity levels grouped by Low and High Water Intake (Hydration Levels). My mean productivity is higher on days with ≥32 ounces (oz) of water intake, which indicates a possible positive relationship between hydration and productivity.
